Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Subjective Wellbeing

### 2021-02-20 Dissonance_Santiago_v3/Dissonance_Santiago_v3.Rmd
Created README.Rmd and updated data package prior to release

@@

load("Dissonance_Santiago.RData")
```

@@

Please note that this document  was prepared using R Markdown and is an example of reproducible research [@Brunsdon2020opening]. The R Markdown file, along with the data file needed to reproduce the analysis, are available for download\footnote{\url{https://drive.google.com/open?id=189ZvfVvRis5xZA9IlviKCLW2bwbPJAgg}}.

>>>>

```{r install-data-package, include=FALSE}
# Run only once if needed
devtools::install_github("paezha/Mode-Dissonance-and-Subjective-Wellbeing", subdir = "ModeDissonanceSantiago")
```

```{r}
library(ModeDissonanceSantiago)
```

@@

#load("Dissonance_Santiago.RData")
```

@@

Please note that this document  was prepared using R Markdown and is an example of reproducible research [@Brunsdon2020opening]. The R Markdown file, along with the data file needed to reproduce the analysis, are available for download\footnote{\url{https://github.com/paezha/Mode-Dissonance-and-Subjective-Wellbeing}}.

### 2021-02-20 README.Rmd
Created README.Rmd and updated data package prior to release

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Affective Values
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
   *Introduction*  
   Subjective wellbeing is a topic that has attracted considerable attention in the transportation literature in recent years. As a result, there is a burgeoning literature that investigates the impacts of travel on subjective wellbeing, and how wellbeing, in turn, can influence behavior. An important aspect of subjective wellbeing are the affective reactions of people to their experiences.   
   *Objective*  
   The objective of this paper is to analyze the affective reactions of travelers with respect to various modes of transportation. In particular, we are interested in the potential for dissonance between primary mode of travel and the mode(s) of travel identified as evoking various affective reactions.  
   *Materials and Methods*  
   The study is based on data collected from a sample of travelers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name their ideal mode(s) of transportation from the perspective of various affective reactions. The reactions we investigate are associated with the values of Freedom, Enjoyment, Happiness, Poverty, Luxury, and Status. Analysis is based on tests of independence and visualization techniques.  
   *Results*  
   The results indicate that users of public transportation experience the most dissonance in terms of affective reactions, and active travelers the least. For those travelers who experience dissonance, active travel is the mode most commonly associated with Freedom, Enjoyment, and Happiness, while public transportation is most commonly associated with Poverty. The automobile, in contrast, is the mode most commonly associated with Luxury and Status.

journal: "Journal of Transport & Health"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
linenumbers: true
numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article

header-includes:
   - \usepackage[margin=1in]{geometry}
   - \usepackage{lineno}
   - \linenumbers
---

<!-- This paper is an R markdown document that includes self-contained, reproducible analysis. To reproduce the paper the following files are needed:

1. Dissonance_Santiago_v1.Rmd (this R markdown document)

2. `Dissonance_Santiago.RData`. For reference: this data file was created using R notebook `00. Analysis of Dissonance Detailed.Rmd`. The initial analysis was done in the same R notebook. Neither of these two notebooks are needed for this document.

3. For compiling into a pdf in the style of an Elsevier document: elsarticle.cls, elsevier-harvard.csl, elsevier-harvard-without-titles.csl, elsevier-vancouver.csl, elsevier-without-titles.cls, elsevier-with-titles.csl, elsevier-with-titles-alphabetical.csl, mybibfile.bib
-->

@@

#load("Dissonance_Santiago.RData")
```

@@

Overall_dissonance <- master %>%
  transmute(USED, 
            FREEDOM_DISSONANCE = ifelse(USED == FREEDOM, 1, 2),
            ENJOYMENT_DISSONANCE = ifelse(USED == ENJOYMENT, 1, 2),
            HAPPINESS_DISSONANCE = ifelse(USED == HAPPINESS, 1, 2),
            POVERTY_DISSONANCE = ifelse(USED == POVERTY, 1, 2),
            LUXURY_DISSONANCE = ifelse(USED == LUXURY, 1, 2),
            STATUS_DISSONANCE = ifelse(USED == STATUS, 1, 2),
            SAFETY_DISSONANCE = ifelse(USED == SAFETY, 1, 2),
            TIME_DISSONANCE = ifelse(USED == TIME, 1, 2),
            EFFICIENCY_DISSONANCE = ifelse(USED == EFFICIENCY, 1, 2),
            AGE,
            EDUCATION,
            INCOME,
            TRAVEL_TIME)

@@

Modes_by_Attribute <- master %>% 
  mutate(FREEDOM = strsplit(as.character(FREEDOM), ""),
         ENJOYMENT = strsplit(as.character(ENJOYMENT), ""),
         HAPPINESS = strsplit(as.character(HAPPINESS), ""),
         POVERTY = strsplit(as.character(POVERTY), ""),
         LUXURY = strsplit(as.character(LUXURY), ""),
         STATUS = strsplit(as.character(STATUS), ""),
         SAFETY = strsplit(as.character(SAFETY), ""),
         TIME = strsplit(as.character(TIME), ""),
         EFFICIENCY = strsplit(as.character(EFFICIENCY), ""))

@@

Please note that this document  was prepared using R Markdown and is an example of reproducible research [@Brunsdon2020opening]. The R Markdown file, along with the data file needed to reproduce the analysis, are available for download\footnote{\url{https://github.com/paezha/Mode-Dissonance-and-Subjective-Wellbeing}}.

@@

In addition, respondents were asked about their primary mode of travel for their regular commute. The modes available were Car; Taxi; _Colectivo_ (a form of shared ride, intermediate in flexibility and capacity between taxi and bus); Motorcycle; Metro; Bus; Bicycle; and Walking. As seen in the top panel of Figure \ref{fig:primary-mode-travel}, the three most common modes of travel are Metro, Bus, and Car, followed by Walking and Bicycle. For the analysis, we aggregate these modes into the following categories (bottom panel of Figure \ref{fig:primary-mode-travel}): (26% of total modes – which corresponds to 91% of the total of private motorized modes), Active (9% - Walking + Bicycle), Public (60% - Metro + Bus), and Other motorized modes (6% - Taxi + Colectivo + Motorcycle).

@@

The premise of this study is that a key component of subjective wellbeing is the affective reaction to experience [see the definition of wellbeing offered by the OECD; -@@oecd2013oecd]. In this paper we investigated mode dissonance from the perspective of six affective values. The research presented here was based on a sample of travelers in Santiago, the capital of Chile. Participants in this research were asked about their typical mode of travel, and then about the mode or modes that they associate with each of the six affective values. Analysis using hypothesis testing (tests of independence) and visualization techniques (bar charts with faceting) uncovered interesting patterns. Some of our findings are well aligned with previous research; for example, active travelers experience less dissonance than car users, and users of public transportation experience the most dissonance of all. However, by considering affective values separately instead of aggregating them into a single indicator of subjective wellbeing, we manage to preserve greater granularity with respect to various responses than most studies. This is important because hedonic/eudaimonic values are more frequently related to active travel (Freedom, Enjoyment, Happiness), while Poverty is more frequently related to public transportation and active travel. Luxury and Status, on the other hand, are more frequently associated to car.

>>>>

---
output: github_document
bibliography: Dissonance_Santiago_v3/mybibfile.bib
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Affective Values

<!-- badges: start -->
<!-- badges: end -->

Beatriz Mella Lira (Pontificia Universidad Católica de Chile)
Antonio Paez (McMaster University)

Journal of Transport and Health (2021), https://doi.org/xx.xxx

## Abstract

### Introduction

Subjective wellbeing is a topic that has attracted considerable attention in the transportation literature in recent years. As a result, there is a burgeoning literature that investigates the impacts of travel on subjective wellbeing, and how wellbeing, in turn, can influence behavior. An important aspect of subjective wellbeing are the affective reactions of people to their experiences.

### Objective

The objective of this paper is to analyze the affective reactions of travelers with respect to various modes of transportation. In particular, we are interested in the potential for dissonance between primary mode of travel and the mode(s) of travel identified as evoking various affective reactions.

### Materials and Methods

The study is based on data collected from a sample of travelers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name their ideal mode(s) of transportation from the perspective of various affective reactions. The reactions we investigate are associated with the values of Freedom, Enjoyment, Happiness, Poverty, Luxury, and Status. Analysis is based on tests of independence and visualization techniques.

### Results

The results indicate that users of public transportation experience the most dissonance in terms of affective reactions, and active travelers the least. For those travelers who experience dissonance, active travel is the mode most commonly associated with Freedom, Enjoyment, and Happiness, while public transportation is most commonly associated with Poverty. The automobile, in contrast, is the mode most commonly associated with Luxury and Status.

## Keywords

- Mode of travel  
- Subjective wellbeing  
- Dissonance  
- Affective values  
- Santiago de Chile  
- Survey  
- Data analysis

@@

#load("Dissonance_Santiago_v3/Dissonance_Santiago.RData")
```

@@

Overall_dissonance <- Dissonance_Santiago %>%
  transmute(USED, 
            FREEDOM_DISSONANCE = ifelse(USED == FREEDOM, 1, 2),
            ENJOYMENT_DISSONANCE = ifelse(USED == ENJOYMENT, 1, 2),
            HAPPINESS_DISSONANCE = ifelse(USED == HAPPINESS, 1, 2),
            POVERTY_DISSONANCE = ifelse(USED == POVERTY, 1, 2),
            LUXURY_DISSONANCE = ifelse(USED == LUXURY, 1, 2),
            STATUS_DISSONANCE = ifelse(USED == STATUS, 1, 2),
            SAFETY_DISSONANCE = ifelse(USED == SAFETY, 1, 2),
            TIME_DISSONANCE = ifelse(USED == TIME, 1, 2),
            EFFICIENCY_DISSONANCE = ifelse(USED == EFFICIENCY, 1, 2),
            AGE,
            EDUCATION,
            INCOME,
            TRAVEL_TIME)

@@

Modes_by_Attribute <- Dissonance_Santiago %>% 
  mutate(FREEDOM = strsplit(as.character(FREEDOM), ""),
         ENJOYMENT = strsplit(as.character(ENJOYMENT), ""),
         HAPPINESS = strsplit(as.character(HAPPINESS), ""),
         POVERTY = strsplit(as.character(POVERTY), ""),
         LUXURY = strsplit(as.character(LUXURY), ""),
         STATUS = strsplit(as.character(STATUS), ""),
         SAFETY = strsplit(as.character(SAFETY), ""),
         TIME = strsplit(as.character(TIME), ""),
         EFFICIENCY = strsplit(as.character(EFFICIENCY), ""))

@@

Please note that this document  was prepared using R Markdown and is an example of reproducible research [@Brunsdon2020opening]. The R Markdown file, along with the data file needed to reproduce the analysis, are available for download [here](https://github.com/paezha/Mode-Dissonance-and-Subjective-Wellbeing).

@@

In addition, respondents were asked about their primary mode of travel for their regular commute. The modes available were Car; Taxi; _Colectivo_ (a form of shared ride, intermediate in flexibility and capacity between taxi and bus); Motorcycle; Metro; Bus; Bicycle; and Walking. As seen in the top panel of Figure \ref{fig:primary-mode-travel}, the three most common modes of travel are Metro, Bus, and Car, followed by Walking and Bicycle. For the analysis, we aggregate these modes into the following categories (bottom panel of Figure \ref{fig:primary-mode-travel}): (26% of total modes – which corresponds to 91% of the total of private motorized modes), Active (9% \- Walking + Bicycle), Public (60% \- Metro + Bus), and Other motorized modes (6% \- Taxi + Colectivo + Motorcycle).

@@

The premise of this study is that a key component of subjective wellbeing is the affective reaction to experience [see the definition of wellbeing offered by the OECD; -@oecd2013oecd]. In this paper we investigated mode dissonance from the perspective of six affective values. The research presented here was based on a sample of travelers in Santiago, the capital of Chile. Participants in this research were asked about their typical mode of travel, and then about the mode or modes that they associate with each of the six affective values. Analysis using hypothesis testing (tests of independence) and visualization techniques (bar charts with faceting) uncovered interesting patterns. Some of our findings are well aligned with previous research; for example, active travelers experience less dissonance than car users, and users of public transportation experience the most dissonance of all. However, by considering affective values separately instead of aggregating them into a single indicator of subjective wellbeing, we manage to preserve greater granularity with respect to various responses than most studies. This is important because hedonic/eudaimonic values are more frequently related to active travel (Freedom, Enjoyment, Happiness), while Poverty is more frequently related to public transportation and active travel. Luxury and Status, on the other hand, are more frequently associated to car.
